---
title: "Engage-Integrate"
author: "Your Name"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background on the data set

a)

https://www.kaggle.com/datasets/maharshipandya/-spotify-tracks-dataset

b) 
This data set contains tracks/songs found on Spotify, a music streaming platform. The list was created to include 125 distinct genres of music. The process of picking the tracks is unknown. An observation in this data set holds information about a specific track/song. The variables include the Spotify track ID, artists included on the track, album that the track came from, the name of the track, popularity ranked between 0-100, duration of the track, genre, and other numerical variables (such as danceability, key, and instrumentalness) that describe the sound produced from the track. Danceability describes shows "danceable" a track is based on the combined values of tempo, rhythm, beat strength, and other sound components. 



```{r}
library(tidyverse)
library(dplyr)
tracks<- read_csv("Spotify Tracks Dataset.csv")
glimpse(tracks)
```

d) **Representing single variables**

```{r}
tracks %>% 
  ggplot(aes(x = popularity)) + geom_histogram()

```

```{r}
tracks %>% 
  ggplot(aes(x = duration_ms)) + geom_boxplot()
```

```{r}
tracks %>% 
  group_by(track_genre) %>% 
  ggplot(aes(x = track_genre)) + geom_bar()
```

e) 
```{r}
tracks %>% 
  filter(popularity == 0) %>% 
  count(track_genre) %>% 
  arrange(desc(n))
# I noticed that there were many tracks that had a popularity of 0. I wanted to investigate the genres in this category to see if there was a correlation or pattern. Jazz, Iranian, and Romance were the genres with the highest number of tracks with a popularity of 0. 
```
```{r}
duration_change <- tracks %>%
  filter(duration_ms != 0) %>% 
  mutate(duration_min = duration_ms/60000) %>% 
  arrange(desc(duration_min))

summary(duration_change)

IQR = 4.3584 - 2.9011
4.3584 + 1.5*IQR

duration_change %>% 
  filter(duration_min > 6.54435) %>% 
  group_by(track_genre) %>% 
  count(track_genre) %>% 
  arrange(desc(n))
# From the original plot, there were many extreme outliers on the higher end that I wanted to investigate. I wanted to see if there was a correlation between the genre and the presented outliers. To do this, I calculated the range of the greater outliers and applied that to the data set. The most common genres found in this set are minimal-techno, detroit-techno, and chicago-house. I also converted the duration to minutes so it would be easier for me to understand. 
```

```{r}
tracks %>% 
  count(track_genre)
# The bar plot seemed strange. With further investigation, there are 114 genres and 1000 tracks for each genre. This explains the shape/distribution of the plot.
```

f) **Make a plot that could help you understand whether there is an association between two or more variables.**

```{r}
tracks %>% 
  filter(popularity >= 50 )%>% 
  ggplot(aes(x = acousticness, y = loudness, color = energy)) + geom_point(alpha = 0.1)
```


g) **Linear Regression**

```{r}
library(moderndive)
fit <- lm(danceability ~ tempo * energy + track_genre, data = tracks)

get_regression_table(fit)

```

h) **Interpretation**

The coefficient values are consist with what I expected. Danceability measures how suitable a song is for dancing and is mainly calculated with tempo and energy and can vary within each genre. In this dataset, it is measured between 0 (lowest danceability) and 1 (highest danceability).Genres like death metal have low danceability, while genres like reggaeton have high danceability. As a result, the genres with low danceability have lower coefficients and high danceability have high coefficients.
